MyTitle

### 2024-11-03 paper/main.tex
updated

% 定理
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{dfn}[thm]{Definition}
\newtheorem{cor}[thm]{Corollary}

@@

By choosing $\mathbb{P}$ appropriately, we can construct $M[G]$ with various properties.
What holds or does not hold in $M[G]$ can be identified using the forcing relation.
The forcing relation is defined recursively on formulas in a forcing language.
The forcing language is an extension of tha language of set theory by adding the elements of $M^{\mathbb{P}}$ as constants.
\begin{dfn} %kunen
  We define the \emph{forcing relation} $\Vdash$ for formulas in the forcing language and $p \in \mathbb{P}$ inductively
  \footnote{
    Specifically, the forcing relation is first defined for atomic formulas $\dot{x} = \dot{y}$ and $\dot{x} \in \dot{y}$ by mutual recursion,
    inductively on the $\mathbb{P}$-rank of $\dot{x}$ and $\dot{y}$.
    Then, the definition is extended to all formulas in the forcing language by induction on the complexity of formulas.
  }
  as follows:
  \begin{enumerate}
    \item $p \Vdash \dot{x} = \dot{y} \Leftrightarrow \forall \dot{z} \in \mathrm{dom}(\dot{x}) \cup \mathrm{dom}(\dot{y})\,
\forall q \leq_{\mathbb{P}} p \, (q \Vdash \dot{z} \in \dot{x} \leftrightarrow q \Vdash \dot{z} \in \dot{y})$
    \item $p \Vdash \dot{x} \in \dot{y} \Leftrightarrow \forall q \leq_{\mathbb{P}} p\, \exists r \leq_{\mathbb{P}} q \, 
    \exists s \in \mathbb{P} \, \exists \dot{z} \in M^{\mathbb{P}} \, (\langle \dot{z}, s \rangle \in \dot{y} \land r \leq_{\mathbb{P}} s \land r \Vdash \dot{x} = \dot{z} )$
    \item $p \Vdash \phi \land \psi \Leftrightarrow p \Vdash \phi \land p \Vdash \psi$
    \item $p \Vdash \neg \phi \Leftrightarrow \forall q \leq_{\mathbb{P}} p \, \neg (q \Vdash \phi)$
    \item $p \Vdash \exists x \phi(x) \Leftrightarrow \forall q \leq_{\mathbb{P}} p \, \exists r \leq_{\mathbb{P}} q \, \exists \dot{x} \in M^{\mathbb{P}} \, (r \Vdash \phi(\dot{x}))$
  \end{enumerate}
\end{dfn}

\begin{thm}
  Let $G$ be a generic filter on $\mathbb{P}$, $\varphi$ be a formula, and $\dot{x} \in M^{\mathbb{P}}$, then 
  $$M[G] \vDash \varphi(\dot{x}^G) \Leftrightarrow \exists p \in G \, (p \Vdash \varphi(\dot{x}))$$
\end{thm}

@@

\subsection{Symmetric Extensions}
Symmetric extensions are substructures of generic extensions of a given c.t.m.\  of ZF 
and are formed by interpreting only the hereditarily symmetric names.
Let $M$ be a c.t.m.\  of ZF, $(\mathbb{P}, \leq_{\mathbb{P}})$ be a pre-ordered set in $M$ with the maximum element $1_{\mathbb{P}}$.

@@

\begin{thm}\label{thm:symmetry_main_thm}
  Let $G$ be a generic filter on $\mathbb{P}$, $\mathcal{N}$ be a symmetric extension generated by $G$, $\varphi$ be a formula, and $\dot{x} \in \mathrm{HS}$, then
  $$\mathcal{N} \vDash \varphi(\dot{x}) \Leftrightarrow \exists p \in G \, (p \Vdash_{\mathrm{HS}} \varphi(\dot{x}))$$
\end{thm}

\begin{cor}\label{cor:symmetry_main_cor}
  Let $p \in \mathbb{P}$, $\varphi$ be a formula, and $\dot{x} \in \mathrm{HS}$, then
  $$p \Vdash_{\mathrm{HS}} \varphi(\dot{x}) \Leftrightarrow \text{for any generic filter } G \text{ containing } p, \mathcal{N} \vDash \varphi(\dot{x}^G)$$
  Where $\mathcal{N}$ is the symmetric extension generated by $G$.
\end{cor}

@@

\begin{lem}\label{lem:symmetry_lemma}
  Let $\pi$ be an automorphism of $\mathbb{P}$, $\dot{x}_0, \ldots, \dot{x}_n \in \mathrm{HS}$, and $\varphi$ be a formula, then
  $$p \Vdash_{\mathrm{HS}} \varphi(\dot{x}_0, \ldots, \dot{x}_n) \Leftrightarrow \pi p \Vdash_{\mathrm{HS}} \varphi(\pi \dot{x}_0, \ldots, \pi \dot{x}_n)$$
\end{lem}

@@

\section{Isabelle/ZF and Formalization in Prior Work}
In this section, we introduce Isabelle/ZF, a proof assistant for ZF set theory, 
and the results from prior work used in the formalization of this study.
\subsubsection{The locale \texttt{"forcing\_data"} by Gunther et al.}
\begin{comment}  
\section{Isabelle/ZF}
\section{Internalized First-Order Formulas}
\section{Forcing}  
\end{comment}

@@

\subsubsection{Hereditarily Symmetric Names}
We continue our work within the locale \texttt{"M\_symmetric\_system"}.
We define the set of hereditarily symmetric names \texttt{"HS"} 
\begin{isaframe}
\begin{minted}[breaklines]{isabelle}
definition sym where "sym(x) ≡ { π ∈ 𝒢. Pn_auto(π)`x = x }"

@@

\section{Proving that Symmetric Extensions are Models of ZF}
\begin{comment}
\subsection{Relativized Forcing Relation $\Vdash_{\mathrm{HS}}$}
\subsection{The Symmetric Lemma}
\subsection{Separation}
\subsection{Replacement}
\subsection{Other Axioms}
\end{comment}

@@

\section{Proving that the Basic Cohen Model satisfies $\neg$AC}

>>>>

% 定理
\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{dfn}[thm]{Definition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{prop}[thm]{Proposition}

@@

By choosing $\mathbb{P}$ appropriately, we can construct $M[G]$ with various properties.
What holds or does not hold in $M[G]$ can be identified using the forcing relation.
The forcing relation is defined recursively on formulas in a forcing language.
The forcing language is an extension of tha language of set theory by adding the elements of $M^{\mathbb{P}}$ as constants.
\begin{dfn} \label{dfn:forcing_relation}
  We define the \emph{forcing relation} $\Vdash$ for formulas in the forcing language and $p \in \mathbb{P}$ inductively
  \footnote{
    Specifically, the forcing relation is first defined for atomic formulas $\dot{x} = \dot{y}$ and $\dot{x} \in \dot{y}$ by mutual recursion,
    inductively on the $\mathbb{P}$-rank of $\dot{x}$ and $\dot{y}$.
    Then, the definition is extended to all formulas in the forcing language by induction on the complexity of formulas.
  }
  as follows:
  \begin{enumerate}
    \item $p \Vdash \dot{x} = \dot{y} \Leftrightarrow \forall \dot{z} \in \mathrm{dom}(\dot{x}) \cup \mathrm{dom}(\dot{y})\,
\forall q \leq_{\mathbb{P}} p \, (q \Vdash \dot{z} \in \dot{x} \leftrightarrow q \Vdash \dot{z} \in \dot{y})$
    \item $p \Vdash \dot{x} \in \dot{y} \Leftrightarrow \forall q \leq_{\mathbb{P}} p\, \exists r \leq_{\mathbb{P}} q \, 
    \exists s \in \mathbb{P} \, \exists \dot{z} \in M^{\mathbb{P}} \, (\langle \dot{z}, s \rangle \in \dot{y} \land r \leq_{\mathbb{P}} s \land r \Vdash \dot{x} = \dot{z} )$
    \item $p \Vdash \phi \land \psi \Leftrightarrow p \Vdash \phi \land p \Vdash \psi$
    \item $p \Vdash \neg \phi \Leftrightarrow \forall q \leq_{\mathbb{P}} p \, \neg (q \Vdash \phi)$
    \item $p \Vdash \exists x \phi(x) \Leftrightarrow \forall q \leq_{\mathbb{P}} p \, \exists r \leq_{\mathbb{P}} q \, \exists \dot{x} \in M^{\mathbb{P}} \, (r \Vdash \phi(\dot{x}))$
  \end{enumerate}
\end{dfn}

\begin{thm}[The Truth Lemma] \label{thm:truth_lemma}
  Let $G$ be a generic filter on $\mathbb{P}$, $\varphi$ be a formula, and $\dot{x} \in M^{\mathbb{P}}$, then 
  $$M[G] \vDash \varphi(\dot{x}^G) \Leftrightarrow \exists p \in G \, (p \Vdash \varphi(\dot{x}))$$
\end{thm}

@@

\subsection{Symmetric Extensions}\label{subsection:symmetrc_extensions}
Symmetric extensions are substructures of generic extensions of a given c.t.m.\  of ZF 
and are formed by interpreting only the hereditarily symmetric names.
Let $M$ be a c.t.m.\  of ZF, $(\mathbb{P}, \leq_{\mathbb{P}})$ be a pre-ordered set in $M$ with the maximum element $1_{\mathbb{P}}$.

@@

\begin{thm}\label{thm:symmetry_truth_lemma}
  Let $G$ be a generic filter on $\mathbb{P}$, $\mathcal{N}$ be a symmetric extension generated by $G$, $\varphi$ be a formula, and $\dot{x} \in \mathrm{HS}$, then
  $$\mathcal{N} \vDash \varphi(\dot{x}) \Leftrightarrow \exists p \in G \, (p \Vdash_{\mathrm{HS}} \varphi(\dot{x}))$$
\end{thm}

\begin{cor}\label{cor:symmetry_truth_lemma_cor}
  Let $p \in \mathbb{P}$, $\varphi$ be a formula, and $\dot{x} \in \mathrm{HS}$, then
  $$p \Vdash_{\mathrm{HS}} \varphi(\dot{x}) \Leftrightarrow \text{for any generic filter } G \text{ containing } p, \mathcal{N} \vDash \varphi(\dot{x}^G)$$
  Where $\mathcal{N}$ is the symmetric extension generated by $G$.
\end{cor}

@@

\begin{lem}[The Symmetry Lemma]\label{lem:symmetry_lemma}
  Let $\pi$ be an automorphism of $\mathbb{P}$, $\dot{x}_0, \ldots, \dot{x}_n \in \mathrm{HS}$, and $\varphi$ be a formula, then
  $$p \Vdash_{\mathrm{HS}} \varphi(\dot{x}_0, \ldots, \dot{x}_n) \Leftrightarrow \pi p \Vdash_{\mathrm{HS}} \varphi(\pi \dot{x}_0, \ldots, \pi \dot{x}_n)$$
\end{lem}

@@

\section{Isabelle/ZF and Formalization in Prior Work}
In this section, we introduce Isabelle/ZF, a proof assistant for ZF set theory, 
and the results from prior work used in the formalization of this study.
\subsubsection{The locale \texttt{"forcing\_data"} by Gunther et al.}

\begin{comment}

強制関係を説明する！

\section{Isabelle/ZF}
\section{Internalized First-Order Formulas}
\section{Forcing}  
\end{comment}

@@

\subsubsection{Hereditarily Symmetric Names}
We continue our work within the locale \texttt{"M\_symmetric\_system"}.
We define the set of hereditarily symmetric names \texttt{"HS"} in the same way as \texttt{"P\_names"}.
This approach makes it easier to use induction based on the $\mathbb{P}$-rank of names.

\begin{isaframe}
\begin{minted}[breaklines]{isabelle}
definition sym where "sym(x) ≡ { π ∈ 𝒢. Pn_auto(π)`x = x }"

@@

\subsubsection{Formulas Representing These Definitions}
In our proofs, we need to use the separation and replacement axioms in M
based on formulas that include the concepts defined in this section.
Therefore, it is necessary to represent these concepts as internalized formulas.

We utilized the automated formula synethisis method by Gunther et al. \cite{gunther_forcing};
however, it could not applied to complex formulas, 
and re-verifying the properties of the derived formulas in an accessible form remained a repititive task.

Example of the formulas we defined includes \texttt{"is\_P\_auto\_group"}, \texttt{"is\_P\_name\_fm"}, \texttt{"is\_Pn\_auto\_fm"}, and \texttt{"is\_HS\_fm"},
whick represent \texttt{"P\_names"}, \texttt{"Pn\_auto"} and \texttt{"HS"} respectively.
The correctness of these formulas is verified by proving the following lemmas.

\begin{isaframe}
\begin{minted}[breaklines]{isabelle}
lemma sats_is_P_name_fm_iff : 
    fixes env i j x 
    assumes "env ∈ list(M)" "i < length(env)" "j < length(env)" 
                    "nth(i, env) = P" "nth(j, env) = x"  
    shows "sats(M, is_P_name_fm(i, j), env) ⟷ x ∈ P_names"

lemma sats_is_Pn_auto_fm_iff :
    fixes x π v env i j k l
    assumes "i < length(env)" "j < length(env)" "k < length(env)" "l < length(env)" 
                    "nth(i, env) = P" "nth(j, env) = π" "nth(k, env) = x" "nth(l, env) = v" 
                    "env ∈ list(M)" "π ∈ P_auto" 
    shows "sats(M, is_Pn_auto_fm(i, j, k, l), env) ⟷ x ∈ P_names ∧ v = Pn_auto(π)`x"

lemma sats_is_HS_fm_iff : 
    fixes x i j env 
    assumes "env ∈ list(M)" "i < length(env)" "j < length(env)" 
                    "nth(i, env) = <ℱ, 𝒢, P, P_auto>" "nth(j, env) = x" 
    shows "sats(M, is_HS_fm(i, j), env) ⟷ x ∈ HS" 
\end{minted}
\end{isaframe}

Note that these formulas take as parameters various elements fixed within the locale, 
such as the notion of forcing \texttt{"P"}, the normal filter \texttt{"ℱ"}, etc.

\section{Proving Symmetric Extensions are Models of ZF}

In this section, we prove that symmetric extensions are models of ZF.
Initially, we attempted to prove this based on its proof in Karagila's lecture note \cite{karagila},
but this approach was unsuccessful.
Instead, we prove this by using the relativized forcing relation $\Vdash_{\mathrm{HS}}$.
Karagila's approach is applicable to a more general situation,
without using forcing or generic and symmetric extensions,
whereas our approach is more specific to the context of symmetric extensions.

\subsubsection{Unsuccessful Approach}

In Karagila's lecture note \cite{karagila},
the fact that symmetric extensions are models of ZF is proven by showing that 
they are internal models of the conrresponding generic extension
generated by the same notion of forcing and generic filter,
using the following proposition:

\begin{prop}
If $\mathcal{N}$ is a transitive class that is almost universal and satisfies $\Delta_0$-separation, 
then $\mathcal{N}$ is a inner model of ZF.
\end{prop}

Where a class $\mathcal{N}$ is said to be \emph{almost universal} 
\footnote{Jech \cite{jech_set_theory}, Theorem 13.9.}
if every subset $X \subseteq \mathcal{N}$ is included in some $Y \in \mathcal{N}$.
Furthermore, $\mathcal{N}$ is said to satisfy $\Delta_0$-separation if it satisfies the separation axioms for formulas with only bounded quantifiers.

We could prove that the symmetric extension \texttt{"SymExt(G)"} is transitive, almost universal and satisfies $\Delta_0$-separation.
However, we could not prove that \texttt{"SymExt(G)"} is actually a class in \texttt{"GenExt(G)"}.
To do this, we needed to explicitly construct a formula that represents the set of all names \texttt{"P\_names"} on the ground model \texttt{"M"} in \texttt{"GenExt(G)"}.
However, we were unable to construct such a formula.

\subsubsection{The Relativized Forcing Relation $\Vdash_{\mathrm{HS}}$}
Instead, we formalize the relativized forcing relation $\Vdash_{\mathrm{HS}}$ and prove that symmetric extensions are models of ZF using it.
This relation is defined by replacing all quantifiers restricted to $M^{\mathbb{P}}$
in the definition of the forcing relation with quantifiers restricted to $\mathrm{HS}$.
By modifying the formalization of forcing by Gunther et al. \cite{gunther_forcing}
\footnote{For non-atomic formulas, the definition of the forcing relation by Gunther et al. \cite{gunther_forcing} 
actually differs from Definition \ref{dfn:forcing_relation}.
In their definition, the unique logical connective is NAND, and the unique quantifier is $\forall$.
Additionally, in the definition, quantifiers restricted to $M$ appear instead of those restricted to $M^{\mathbb{P}}$.}
, we were able to define this relation and prove its properties relatively easily.

Within the locale \texttt{"M\_symmetric\_system"},
we define this relation in the form \texttt{"p ⊩HS φ env"},
which has the same form as the formalization of the original forcing relation \texttt{"p ⊩ φ env"} by Gunther et al.
Our definition satisfies the required properties, such as Theorem \ref{thm:symmetry_truth_lemma}, Corollary \ref{cor:symmetry_truth_lemma_cor}, and Lemma \ref{lem:strengthening}.

\begin{isaframe}
\begin{minted}[breaklines]{isabelle}
lemma HS_truth_lemma:
    assumes 
        "φ∈formula" "M_generic(G)"
    shows 
        "⋀env. env∈list(HS) ⟹ arity(φ)≤length(env) ⟹ 
              (∃p∈G. p ⊩HS φ env) ⟷ SymExt(G), map(val(G),env) ⊨ φ"

lemma definition_of_forcing_HS:
    assumes
        "p∈P" "φ∈formula" "env∈list(HS)" "arity(φ)≤length(env)"
    shows
        "(p ⊩HS φ env) ⟷
              (∀G. M_generic(G) ∧ p∈G  ⟶  SymExt(G), map(val(G),env) ⊨ φ)"

lemma HS_strengthening_lemma:
    assumes 
        "p∈P" "φ∈formula" "r∈P" "r≼p"
    shows
        "⋀env. env∈list(M) ⟹ arity(φ)≤length(env) ⟹ p ⊩HS φ env ⟹ r ⊩HS φ env"
\end{minted}
\end{isaframe}

\subsubsection{The Symmetry Lemma}
Additionally, we prove the symmetry lemma (Lemma \ref{lem:symmetry_lemma}).
Similary to the definitions of the forcing relations, this lemma is proven using a complex induction.
First, we prove this lemma for atomic formulas in the forcing language, 
and then extend the proof to other formulas by structual induction on the formula.
For atomic formulas $\dot{x} = \dot{y}$ and $\dot{x} \in \dot{y}$, 
the definition is mutually recursive, so the lemma for these formulas is proven simultaneously
by induction on the $\mathbb{P}$-rank of $\dot{x}$ and $\dot{y}$.
Specifically, we show these induction steps:

\begin{isaframe}
\begin{minted}[breaklines]{isabelle}
  have MEM_step: "⋀a. Ord(a) ⟹ ∀b ∈ a. Q(EQ, b, b) ⟹ ∀b ∈ a. Q(MEM, b, a)"  
  have EQ_step : "⋀a. Ord(a) ⟹ ∀b ∈ a. Q(MEM, b, a) ⟹ Q(EQ, a, a)"   
\end{minted}
\end{isaframe}

Where \texttt{"Q(MEM, α, β)"} and \texttt{"Q(EQ, α, β)"} are predicates 
representing that the symmetriy lemma holds for $\dot{x} \in \dot{y}$ and $\dot{x} = \dot{y}$, respectively,
for any $\dot{x}$ with $\mathbb{P}$-rank at most $\alpha$ and $\dot{y}$ with $\mathbb{P}$-rank at most $\beta$.

The proof for non-atomic formulas is relatively straightforward, and finally, 
this completes the proof of the lemma.

\begin{isaframe}
\begin{minted}[breaklines]{isabelle}
lemma symmetry_lemma:
    fixes φ π  
    assumes "φ ∈ formula" "is_P_auto(π)" "π ∈ 𝒢" 
    shows "⋀env p. env ∈ list(HS) ⟹ arity(φ) ≤ length(env) ⟹ p ∈ P ⟹ 
                p ⊩HS φ env ⟷ π`p ⊩HS φ map(λx. Pn_auto(π)`x, env)" 
\end{minted}
\end{isaframe}

\subsubsection{Separation}

\begin{comment}
\subsection{Relativized Forcing Relation $\Vdash_{\mathrm{HS}}$}
\subsection{The Symmetric Lemma}
\subsection{Separation}
\subsection{Replacement}
\subsection{Other Axioms}
\end{comment}

@@

\section{Proving the Basic Cohen Model satisfies $\neg$AC}
